---
title: "VarComp"
output:
  html_document: default
  pdf_document: default
date: "2025-04-29"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(matlib)
library(cubature)
library(latex2exp)
library(ggplot2)
library(MCMCpack)
library(coda)
```

#Variance Component Model Gibbs Sampler

The following is the data of interest. The rows represent the batches of which there are 6 in total. Each batch contains 5 elements.
```{r}
# Matrix giving the famous "dyestuff" batch data from Davies (1967).
# Defined so Ydye[i,j] equals yield (in grams) from j'th sample of i'th batch.
# Valid for i=1,2,3,4,5,6, and j=1,2,3,4,5, i.e. I=6 and J=5.

Ydye = t( matrix(
	c(1545, 1440, 1440, 1520, 1580,
         1540, 1555, 1490, 1560, 1495,
         1595, 1550, 1605, 1510, 1560,
         1445, 1440, 1595, 1465, 1545,
         1595, 1630, 1515, 1635, 1625,
         1520, 1455, 1450, 1480, 1445), nrow=5) )

J <- dim(Ydye)[2] # number of samples in each batch
I <- dim(Ydye)[1] # number of batches
```

The following code runs a Gibbs sampler for the model.

```{r}
gibbs_update_u <- function(x, z = rnorm(1,0,1)){
  a <- (a3*x[1]+b3*sum(x[4:(3+I)]))/(x[1]+I*b3)
  b <- sqrt(b3*x[1]/(x[1] + I*b3))
  x[3] <- a + b*z
  return(x)
}

gibbs_update_theta <- function(x, z = rnorm(I,0,1)){
  a <- (x[3]/x[1]+apply(Ydye, MARGIN = 1, FUN = sum)/x[2])/(1/x[1]+J/x[2])
  b <- sqrt(1/(1/x[1] + J/x[2]))
  b <- sqrt(x[1]*x[2]/(x[2] + J*x[1]))
  x[4:(3+I)]  <- a + b * z
  return(x)
}

gibbs_update_V <- function(x, g = rgamma(1, shape = a1+I/2, rate = 1)){
  b <- b1 +sum((x[4:(3+I)]-x[3])^2)/2
  x[1]  <- b/g
  return(x)
}

gibbs_update_W <- function(x, g = rgamma(1, shape = a2+I*J/2, rate = 1)){
  b <- b2 +sum((Ydye-x[4:(3+I)])^2)/2
  x[2]  <- b/g
  return(x)
}
```

Generate the initial value of $Y_0\sim \nu$
```{r}
par <- c(6, 6, 6, 6, 1600, 1600)

a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
```


```{r}
# M <- 10000 # number of iterations 
# 
# X <- c(100,200,900,rep(900,I))
# X_init <- X
# 
# V <- rinvgamma(1, a1, b1-1)
# W <- rinvgamma(1, a2, b2-1)
# u <- rnorm(1, a3, sqrt(b3))
# theta <- rnorm(I, apply(Ydye,2,mean), W/J)
# Y <- c(V[1], W[1], u[1], theta)
# Y_init <- Y
# 
# df_X <- data.frame(t(c(X,0,0)))
# df_Y <- data.frame(t(c(Y,0,0)))
# colnames(df_X) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
# colnames(df_Y) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
# 
# #a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
#   
# for (i in 1:M) {
#     for (coord in 1:4){
#       if (coord == 3) { # Modify coordinate u
#         z = rnorm(1,0,1)
#         X <- gibbs_update_u(X, z)
#         Y <- gibbs_update_u(Y, z)
#       }
#       else if (coord == 4) { # Modify all K thetas
#         z = rnorm(I,0,1)
#         X <- gibbs_update_theta(X, z)
#         Y <- gibbs_update_theta(Y, z)
#       }
#       else if (coord == 1) { # Modify coordinate V
#         g = rgamma(1, shape = a1+I/2, rate = 1)
#         X <- gibbs_update_V(X, g)
#         Y <- gibbs_update_V(Y, g)
#       }
#       else if (coord == 2) { # Modify coordinate W
#         g = rgamma(1, shape = a2+I*J/2, rate = 1)
#         X <- gibbs_update_W(X,g)
#         Y <- gibbs_update_W(Y,g)
#       }
#       df_X <- rbind(df_X,c(X, coord, i))
#       df_Y <- rbind(df_Y,c(Y, coord, i))
#     }
# }
```

```{r}
# save(df_X, file = "df_X_VarComp.RData")
# save(df_Y, file = "df_Y_VarComp.RData")
```

```{r}
load('df_X_VarComp.RData')
load('df_Y_VarComp.RData')
```

```{r}
df_X %>%
  filter(iter>9000, coord==1) %>% 
  ggplot(aes(x = iter, y = theta1)) +
  geom_line() + labs(title="theta1") + theme_bw()

df_X %>%
  filter(iter>9000, coord==1) %>% 
  ggplot(aes(x = iter, y = mu)) +
  geom_line() + labs(title="mu") + theme_bw()

df_X %>%
  filter(iter>9000, coord==1) %>% 
  ggplot(aes(x = iter, y = V)) +
  geom_line() + labs(title="V") + theme_bw()

df_X %>%
  filter(iter>9000, coord==1) %>% 
  ggplot(aes(x = iter, y = W)) +
  geom_line() + labs(title="W") + theme_bw()
```

```{r}
df_X %>%
  filter(iter>25, coord==1) %>% 
  ggplot(aes(x = iter, y = theta1)) +
  geom_line() + labs(title="theta1") + theme_bw()

df_X %>%
  filter(iter>25, coord==1) %>% 
  ggplot(aes(x = iter, y = mu)) +
  geom_line() + labs(title="mu") + theme_bw()

df_X %>%
  filter(iter>25, coord==1) %>% 
  ggplot(aes(x = iter, y = V)) +
  geom_line() + labs(title="V")+ theme_bw()

df_X %>%
  filter(iter>25, coord==1) %>% 
  ggplot(aes(x = iter, y = W)) + 
  geom_line() + labs(title="W") + theme_bw()
```

```{r}
df_X1 <- df_X %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "X")
df_Y1 <- df_Y %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "Y")

df <- left_join(df_X1, df_Y1, by=c('coord','iter','names'))
df <- df %>% mutate(diff = X-Y)

df_diff <- df %>%
  filter(coord ==1) %>%
  group_by(iter) %>%
  summarise(diff = sum(abs(diff)))
```


```{r}
df_diff %>%
  filter(iter>2) %>%
  ggplot(aes(x = iter, y = diff)) +
  geom_line() + labs(title = "Value of |X-Y|", subtitle = "Using common random number technique") + theme_bw() +
  xlab("iteration") + ylab("value")
```

```{r}
#par <- c(6, 6, 6, 6, 1600, 1600)

simVCM <- function(sim_no){
  a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
  
  V <- rinvgamma(1, a1, b1-1)
  W <- rinvgamma(1, a2, b2-1)
  u <- rnorm(1, a3, sqrt(b3))
  theta <- rnorm(I, apply(Ydye,2,mean), W/J)
  Y <- c(V[1], W[1], u[1], theta)
  Y_init <- Y

  M <- 10000
  X <- c(100,200,900,rep(900,I))
  X_init <- X
  
  df_X <- data.frame(t(c(X,0,0)))
  df_Y <- data.frame(t(c(Y,0,0)))
  colnames(df_X) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
  colnames(df_Y) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
  
  a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
    
  for (i in 1:M) {
      for (coord in 1:4){
        if (coord == 3) { # Modify coordinate u
          z = rnorm(1,0,1)
          X <- gibbs_update_u(X, z)
          Y <- gibbs_update_u(Y, z)
        }
        else if (coord == 4) { # Modify all I thetas
          z = rnorm(I,0,1)
          X <- gibbs_update_theta(X, z)
          Y <- gibbs_update_theta(Y, z)
        }
        else if (coord == 1) { # Modify coordinate V
          g = rgamma(1, shape = a1+I/2, rate = 1)
          X <- gibbs_update_V(X, g)
          Y <- gibbs_update_V(Y, g)
        }
        else if (coord == 2) { # Modify coordinate W
          g = rgamma(1, shape = a2+I*J/2, rate = 1)
          X <- gibbs_update_W(X,g)
          Y <- gibbs_update_W(Y,g)
        }
        df_X <- rbind(df_X,c(X, coord, i))
        df_Y <- rbind(df_Y,c(Y, coord, i))
      }
  }

  df_X1 <- df_X %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "X")
  df_Y1 <- df_Y %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "Y")
  
  df <- left_join(df_X1, df_Y1, by=c('coord','iter','names'))
  df <- df %>% mutate(diff = X-Y)
  
  df_diff <- df %>%
    filter(coord ==1) %>%
    group_by(iter) %>%
    summarise(diff = sum(abs(diff))) %>% 
    mutate(sim_no=as.character(sim_no))
  return(df_diff)
}
```

```{r}
# df_diff_all <- simVCM(1)
# for(i in 2:300){
#   df <- simVCM(i)
#   df_diff_all <- rbind(df_diff_all, df)
#   print(i)
# }
```

```{r}
# save(df_diff_all, file = "df_diff_all_VarComp.RData")
```


```{r}
load('df_diff_all_VarComp.RData')
```

```{r}
df_diff_all %>%
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), min_diff =  min(diff), max_diff = max(diff)) %>% 
  filter(iter>2) %>%  
  ggplot(aes(x = iter)) +
  geom_ribbon(aes(ymin = min_diff, ymax = max_diff), fill = 'grey', alpha=0.5) +
  geom_line(aes(y = mean_diff)) +
  theme_bw() +
  #labs(title = TeX("Value of $||X_n-X'_n||_1$"), subtitle = "Using common random number technique") +
  xlab(TeX("n=iteration")) + ylab(TeX("$||X_n-X'_n||_1$"))
```



Calculate $K$ for the variance component model.

## Find the constant $K$ from theorem 4.4


Step 1: Find an upper bound on $f/\nu$


```{r}
a_w <- I*(J-1)/2-1
b_w <- 1
a_v <- I/2-1
b_v <- 1

V_star <- b_v/(a_v+1)
W_star <- b_w/(a_w+1)

C <- (b1/(b1-1))^a1*(b2/(b2-1))^a2*1/((2*pi)^(I*J/2)*J^(1/2))

sup_ratio <- C*gamma(a_v)/(b_v^a_v)*dinvgamma(V_star,a_v,b_v)*
  gamma(a_w)/(b_w^a_w)*dinvgamma(W_star,a_w,b_w)
sup_ratio <- C*(exp(-b_v/V_star)/(V_star)^(a_v+1))*(exp(-b_w/W_star)/(W_star)^(a_w+1))
```


Step 2: Find a lower bound on the unnormalized distribution function, $L$.


```{r}
dlognorm <- function(x, m, sd){
  return(-0.5*log(2*pi*sd^2)-(x-m)^2/(2*sd^2))
}

dloginvgamma <- function(x, a, b){
  return(a*log(b)-log(gamma(a))-(a+1)*log(x)-b/x)
}
```


```{r}
logf <- function(x, Y=Ydye, para=par){
  V <- x[1]; W <- x[2]; mu <- x[3]; theta <- x[4:length(x)]
  if  (V<=0 || W<=0){
    return(0)
  }
  a1 <- para[1]; b1 <- para[2]; a2 <- para[3]; b2 <- para[4]; a3 <- para[5]; b3 <- para[6];
  
  f_v <- dloginvgamma(V, a1, b1)
  f_w <- dloginvgamma(W, a2, b2)
  f_u <- dlognorm(mu, a3, sqrt(b3))
  
  f_theta <- 0
  for (i in 1:length(theta)){
    f_theta_i <- dlognorm(theta[i], mu, sqrt(V))
    f_theta <- f_theta + f_theta_i
  }
  
  f_y <- 0
  for(i in 1:length(theta)){
    for(j in 1:length(Ydye[i,])){
      f_y_i <- dlognorm(Ydye[i,j],theta[i],sqrt(W))
      f_y <- f_y + f_y_i
    }
  }
  
  logf <- f_v+f_w+f_u+f_theta+f_y
  if (logf == -Inf){
    return(0)
  }
  else{
    return(logf) 
  }
  
}

f <- function(x, Y=Ydye, para=par){
  logf_val <- logf(x, Y, para)
if (logf_val == 0){
    return(0)
  }
  else{
    return(exp(logf(x, Y, para)))
  }
}
```


```{r}
x <- c(1.1,3000,1520,1520,1520,1520,1520,1520,1520)

int1 <- adaptIntegrate(logf, lowerLimit = x-rep(0.5, length(x)), upperLimit = x-rep(0.0, length(x)))

int1_val <- int1$integral-int1$error

L=exp(int1_val)
L
```
According to equation 15, an upper bound on K is as follows,
```{r}
K <- log(sup_ratio)-int1_val
K <- exp(K)
K
```
```{r}
df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), min_diff =  min(diff), max_diff = max(diff),lq_diff =  quantile(diff, probs = c(0.25)), uq_diff = quantile(diff, probs = c(0.75))) %>% 
  filter(iter>2) %>%  
  ggplot(aes(x = iter)) +
  geom_ribbon(aes(ymin = min_diff, ymax = max_diff), fill = 'grey', alpha=0.5) +
  geom_ribbon(aes(ymin = lq_diff, ymax = uq_diff), fill = 'grey', alpha=0.75) +
  geom_line(aes(y = mean_diff), size=1) + scale_y_log10() +
  theme_bw() + geom_hline(yintercept=0.01) +
  #labs(title = TeX("Value of $||X_n-X'_n||_1$"), subtitle = "Using common random number technique") +
  xlab(TeX("n=iteration")) + ylab(TeX("Logscale of $K||X_n-X'_n||_1$"))
```
```{r}
df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), 
            min_diff =  min(diff), 
            max_diff = max(diff),
            lq0_diff =  quantile(diff, probs = c(0.05)), 
            uq0_diff = quantile(diff, probs = c(0.95)), 
            lq1_diff =  quantile(diff, probs = c(0.1)), 
            uq1_diff = quantile(diff, probs = c(0.9)), 
            lq2_diff =  quantile(diff, probs = c(0.2)), 
            uq2_diff = quantile(diff, probs = c(0.8)), 
            lq3_diff =  quantile(diff, probs = c(0.3)), 
            uq3_diff = quantile(diff, probs = c(0.7)), 
            lq4_diff =  quantile(diff, probs = c(0.4)), 
            uq4_diff = quantile(diff, probs = c(0.6))) %>% 
  filter(iter>2) %>%
  ggplot(aes(x = iter)) +
  geom_ribbon(aes(ymin = min_diff, ymax = max_diff), fill = "skyblue", alpha=0.15) +
  geom_ribbon(aes(ymin = lq0_diff, ymax = uq0_diff), fill = "skyblue", alpha=0.20) +
  geom_ribbon(aes(ymin = lq1_diff, ymax = uq1_diff), fill = "skyblue", alpha=0.40) +
  geom_ribbon(aes(ymin = lq2_diff, ymax = uq2_diff), fill = "skyblue", alpha=0.60) +
  geom_ribbon(aes(ymin = lq3_diff, ymax = uq3_diff), fill = "skyblue", alpha=0.80) +
  geom_ribbon(aes(ymin = lq4_diff, ymax = uq4_diff), fill = "skyblue", alpha=0.90) +
  geom_line(aes(y = mean_diff),size=1) + scale_y_log10() + geom_hline(yintercept=0.01, linetype=2) +
  #labs(title = TeX("Value of $K||X_n-X'_n||_1$"), subtitle = "Using common random number technique") +
  theme_bw(base_size = 13) + scale_y_log10() +
  xlab(TeX("n=iteration")) + ylab(TeX("Logscale of $W(L(X_n),\\pi)$")) +
  geom_text(aes(200,0.003,label = 'y=0.01', vjust = -1)) 
  #scale_alpha_manual(values=c(seq(0.05,0.9,length.out=floor(0.5*length(qq))),
                              #seq(0.9,0.05,length.out=floor(0.5*length(qq)))))
```

```{r}
first_hit_wass <-
  df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff)) %>% 
  filter(mean_diff<=0.01) %>% 
  summarise(min(iter))

first_hit_wass <- first_hit_wass$`min(iter)`

diff9900 <- df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff)) %>% 
  filter(iter == first_hit_wass)

diff9900
  
```

## Comparing this bound to Cowles and Rosenthal's

Generate the initial value of $Y_0\sim \nu$
```{r}
par <- c(0.5,1.00001,0.00001,1.00001, 0, 10^(12))   # initial values from Cowles and Rosenthal. Note that b1 is not 1, but 1.0001. This is because our method requires b1>1.
a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
```

```{r}
# Generate initial values that are aligned with Cowles and Rosenthal
batch_m <- apply(Ydye,1,mean)
bar_y <- mean(Ydye)
v1 <- sum((Ydye-batch_m)^2)/(I*J)
v2 <- sum((batch_m-bar_y)^2)/I

theta <- (J*v1*batch_m +v2*bar_y)/(J*v1+v2)
u <- bar_y
```


```{r}
simVCM <- function(){ #same as previous simVCM function except that X_init is different and W is assigned 0 (since choice of priors would result in Inf)
  M <- 700
  a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
  
  #initial value of mu and theta in Y coincides with V(X_0)=0 (eqn 9 from Cowles and Rosenthal paper). initial values of V and W are taken as conditional distributions given mu and theta.
  V <- rinvgamma(1, a1 + I/2, b1+sum((theta-u)^2))
  W <- rinvgamma(1, a2+I*J/2, b2 + sum((Ydye-theta)^2))
  Y <- c(V[1], W[1], u, theta)
  Y_init <- Y
  
  # initial value of X coincides with eqn 6 from paper
  V <- rinvgamma(1, a1, b1-1)
  W <- rinvgamma(1, a2, 0)
  u <- rnorm(1, a3, sqrt(b3))
  theta <- rnorm(I, apply(Ydye,2,mean), W/J)
  X <- c(V[1], W[1], u[1], theta)
  X_init <- X

  df_X <- data.frame(t(c(X,0,0)))
  df_Y <- data.frame(t(c(Y,0,0)))
  colnames(df_X) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
  colnames(df_Y) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")

 # a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];

  for (i in 1:M) {
      for (coord in 1:4){
        if (coord == 3) { # Modify coordinate u
          z = rnorm(1,0,1)
          X <- gibbs_update_u(X, z)
          Y <- gibbs_update_u(Y, z)
        }
        else if (coord == 4) { # Modify all I thetas
          z = rnorm(I,0,1)
          X <- gibbs_update_theta(X, z)
          Y <- gibbs_update_theta(Y, z)
        }
        else if (coord == 1) { # Modify coordinate V
          g = rgamma(1, shape = a1+I/2, rate = 1)
          X <- gibbs_update_V(X, g)
          Y <- gibbs_update_V(Y, g)
        }
        else if (coord == 2) { # Modify coordinate W
          g = rgamma(1, shape = a2+I*J/2, rate = 1)
          X <- gibbs_update_W(X,g)
          Y <- gibbs_update_W(Y,g)
        }
        df_X <- rbind(df_X,c(X, coord, i))
        df_Y <- rbind(df_Y,c(Y, coord, i))
      }
  }

  df_X1 <- df_X %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "X")
  df_Y1 <- df_Y %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "Y")

  df <- left_join(df_X1, df_Y1, by=c('coord','iter','names'))
  df <- df %>% mutate(diff = X-Y)

  df_diff <- df %>%
    filter(coord ==1) %>%
    group_by(iter) %>%
    summarise(diff = sum(abs(diff)))
  return(df_diff)
}
```

```{r}
# df_diff_all <- simVCM()
# df_diff_all <- df_diff_all %>% mutate(sim_no=1)
# for (i in 2:2000){
#   df_diff <- simVCM()
#   df_diff <- df_diff %>% mutate(sim_no=i)
#   df_diff_all <- rbind(df_diff_all,df_diff)
# }
# 
# save(df_diff_all, file = "df_diff_all_Cowles_VarComp.RData")
```

```{r}
load('df_diff_all_Cowles_VarComp.RData')
```


```{r}
df_diff_all %>%
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), min_diff =  min(diff), max_diff = max(diff)) %>% 
  filter(iter==471)
```


## Find the constant $K$ from theorem 4.4 with priors specified by Cowles
Generating bound between |X_n-Y_n| and |X_n-X_infty|

```{r}
V_star <- 2/I
W_star <- 2/(J*I-I)

C <- (b1/(b1-1))^a1*(b2/(b2-1))^a2*1/((2*pi)^(I*J/2)*J^(1/2))

sup_ratio <- C*(exp(-1/V_star)/(V_star)^(I/2))*(exp(-1/W_star)/(W_star)^((J*I-I)/2))
sup_ratio
```

```{r}
x <- c(1000,3000,1520,1520,1520,1520,1520,1520,1520)

int1 <- adaptIntegrate(logf, lowerLimit = x-rep(0.5, length(x)), upperLimit = x+rep(0.10, length(x)))

int1_val <- int1$integral-int1$error

L=exp(int1_val)
L
```
According to equation 15, an upper bound on K is as follows,
```{r}
K <- log(sup_ratio)-int1_val
K <- exp(K)
K
```
```{r}
# first time wasserstein distance below 0.01
df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff)) %>% 
  filter(mean_diff<0.01) %>% 
  summarise(min(iter))
```


```{r}
df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), 
            min_diff =  min(diff), 
            max_diff = max(diff),
            lq0_diff =  quantile(diff, probs = c(0.05)), 
            uq0_diff = quantile(diff, probs = c(0.95)), 
            lq1_diff =  quantile(diff, probs = c(0.1)), 
            uq1_diff = quantile(diff, probs = c(0.9)), 
            lq2_diff =  quantile(diff, probs = c(0.2)), 
            uq2_diff = quantile(diff, probs = c(0.8)), 
            lq3_diff =  quantile(diff, probs = c(0.3)), 
            uq3_diff = quantile(diff, probs = c(0.7)), 
            lq4_diff =  quantile(diff, probs = c(0.4)), 
            uq4_diff = quantile(diff, probs = c(0.6))) %>% 
  filter(iter>20, iter<550) %>%  
  ggplot(aes(x = iter)) +
  geom_ribbon(aes(ymin = min_diff, ymax = max_diff), fill = "skyblue", alpha=0.15) +
  geom_ribbon(aes(ymin = lq0_diff, ymax = uq0_diff), fill = "skyblue", alpha=0.20) +
  geom_ribbon(aes(ymin = lq1_diff, ymax = uq1_diff), fill = "skyblue", alpha=0.40) +
  geom_ribbon(aes(ymin = lq2_diff, ymax = uq2_diff), fill = "skyblue", alpha=0.60) +
  geom_ribbon(aes(ymin = lq3_diff, ymax = uq3_diff), fill = "skyblue", alpha=0.80) +
  geom_ribbon(aes(ymin = lq4_diff, ymax = uq4_diff), fill = "skyblue", alpha=0.90) +
  geom_line(aes(y = mean_diff),size=1) + scale_y_log10() + geom_hline(yintercept=0.01, linetype=2) +
  #labs(title = TeX("Value of $K||X_n-X'_n||_1$"), subtitle = "Using common random number technique") +
  theme_bw(base_size = 13) +
  xlab(TeX("n=iteration")) + ylab(TeX("Logscale of $W(L(X_n),\\pi)$")) + 
  geom_text(aes(540,0.004,label = 'y=0.01', vjust = -1)) 
  #scale_alpha_manual(values=c(seq(0.05,0.9,length.out=floor(0.5*length(qq))),
                              #seq(0.9,0.05,length.out=floor(0.5*length(qq)))))
```

```{r}
df_diff_all %>%
  mutate(diff=diff*K) %>% 
  group_by(iter) %>%
  summarise(mean_diff = mean(diff), min_diff =  min(diff), max_diff = max(diff),lq_diff =  quantile(diff, probs = c(0.05)), uq_diff = quantile(diff, probs = c(0.95))) %>% 
  filter(iter>50, iter <560) %>%  
  ggplot(aes(x = iter)) +
  #geom_ribbon(aes(ymin = min_diff, ymax = max_diff), fill = "grey70", alpha=0.50) +
  geom_ribbon(aes(ymin = lq_diff, ymax = uq_diff), fill = "grey70", alpha=0.750) +
  geom_line(aes(y = mean_diff),size=1) + geom_hline(yintercept=0.01) +
  #labs(title = TeX("Value of $K||X_n-X'_n||_1$"), subtitle = "Using common random number technique") +
  theme_bw() + scale_fill_viridis_d() +
  xlab(TeX("n=iteration")) + ylab(TeX("$||L(X_n)-\\pi||_1$"))
```


## Generate bound in TV from Wass distance. 

Must get squared distance 
```{r}
simVCM <- function(){ #same as previous simVCM function except that X_init is different and W is assigned 0 (since choice of priors would result in Inf)
  M <- 700
  a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];
  
  #initial value of mu and theta in Y coincides with V(X_0)=0 (eqn 9 from Cowles and Rosenthal paper). initial values of V and W are taken as conditional distributions given mu and theta.
  V <- rinvgamma(1, a1 + I/2, b1+sum((theta-u)^2))
  W <- rinvgamma(1, a2+I*J/2, b2 + sum((Ydye-theta)^2))
  Y <- c(V[1], W[1], u, theta)
  Y_init <- Y
  
  # initial value of X coincides with eqn 6 from paper
  V <- rinvgamma(1, a1, b1-1)
  W <- rinvgamma(1, a2, 0)
  u <- rnorm(1, a3, sqrt(b3))
  theta <- rnorm(I, apply(Ydye,2,mean), W/J)
  X <- c(V[1], W[1], u[1], theta)
  X_init <- X

  df_X <- data.frame(t(c(X,0,0)))
  df_Y <- data.frame(t(c(Y,0,0)))
  colnames(df_X) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")
  colnames(df_Y) <- c("V","W","mu",paste("theta",1:I, sep=""),"coord","iter")

 # a1 <- par[1]; b1 <- par[2]; a2 <- par[3]; b2 <- par[4]; a3 <- par[5]; b3 <- par[6];

  for (i in 1:M) {
      for (coord in 1:4){
        if (coord == 3) { # Modify coordinate u
          z = rnorm(1,0,1)
          X <- gibbs_update_u(X, z)
          Y <- gibbs_update_u(Y, z)
        }
        else if (coord == 4) { # Modify all I thetas
          z = rnorm(I,0,1)
          X <- gibbs_update_theta(X, z)
          Y <- gibbs_update_theta(Y, z)
        }
        else if (coord == 1) { # Modify coordinate V
          g = rgamma(1, shape = a1+I/2, rate = 1)
          X <- gibbs_update_V(X, g)
          Y <- gibbs_update_V(Y, g)
        }
        else if (coord == 2) { # Modify coordinate W
          g = rgamma(1, shape = a2+I*J/2, rate = 1)
          X <- gibbs_update_W(X,g)
          Y <- gibbs_update_W(Y,g)
        }
        df_X <- rbind(df_X,c(X, coord, i))
        df_Y <- rbind(df_Y,c(Y, coord, i))
      }
  }

  df_X1 <- df_X %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "X")
  df_Y1 <- df_Y %>% pivot_longer(cols = V:theta6, names_to = "names", values_to = "Y")

  df <- left_join(df_X1, df_Y1, by=c('coord','iter','names'))
  df <- df %>% mutate(diff = X-Y)

  df_diff <- df %>%
    filter(coord ==1) %>%
    group_by(iter) %>%
    summarise(diff = sum((diff)^2))
  return(df_diff)
}
```

```{r}
# df_diff_all2 <- simVCM()
# df_diff_all2 <- df_diff_all2 %>% mutate(sim_no=1)
# for (i in 2:2000){
#   df_diff <- simVCM()
#   df_diff <- df_diff %>% mutate(sim_no=i)
#   df_diff_all2 <- rbind(df_diff_all2,df_diff)
# }
# 
# save(df_diff_all2, file = "df_diff_all2_Cowles_VarComp.RData")
```

```{r}
load('df_diff_all2_Cowles_VarComp.RData')
```

```{r}
df_diff_all2 <- df_diff_all2 %>% 
  mutate(diff = sqrt(diff))
```


```{r}
# I <- dim(Ydye)[2]
# J <- dim(Ydye)[1]

c1=(2/I)*(I/2+a1)*(1+sqrt(2/b1)*1/I+1/(2*b1*I^2))^(I/2+a1-1)*(sqrt(2/b1)+1/(b1*I))

c2=2/(I*J^(3/2))*((I*J)/2+a2)*(1+2/(sqrt(b2)*I*J)+1/(b2*I^2*J^2))^(I*J/2+a2-1)*(2*sqrt(J/b2)+2/(b2*sqrt(J)*I))

TV <- (c1+c2)*J^(1.5)*I
TV
K*TV
```
Add Cowles and Rosenthal's bound
```{r}
x <- 1:10000
DnMBound <- 0.72^(as.integer(0.000152*x))+0.7905*(0.99985)^x
# for(i in 1:700){
#   DnMBound[i] <- 0.072^(as.integer(0.000152*i))+0.7905*(0.99985)^i
# }
tail(DnMBound)
```

```{r}
diffTV <- df_diff_all %>%
  group_by(iter) %>%
  summarise(CRN = mean(diff)*K*TV)
```
```{r}
diffTV_all <- rbind(diffTV, data.frame(iter=701:10000, CRN=rep(0,10000-700))) %>% 
  left_join(data.frame(iter = 1:10000, DnM = DnMBound), by='iter') %>% 
  pivot_longer(cols = c('CRN', 'DnM'), names_to='bound', values_to='val_type') 
# %>% 
#   group_by(iter, bound_type) %>% 
#   mutate(val_type=min(1,val_type))
```
```{r}
diffTV_all 
```


```{r}
diffTV_all %>% 
  filter(iter>470, iter<1000) %>%  
  ggplot(aes(x = iter, y = val_type, colour=bound)) +
  geom_line(aes(lty=bound), size=1) +
  xlab(TeX("n=iteration")) + ylab(TeX("$||L(X_n)-\\pi||_{TV}$")) +
  theme_bw(base_size=13) + theme(legend.position="bottom") + 
  scale_color_manual(values = c('black','skyblue'))
```

```{r}
first_hit_wass <- diffTV %>% 
  filter(CRN <= 0.01) %>% 
  summarise(first_hit_wass=min(iter))

first_hit_wass <- first_hit_wass$first_hit_wass

diffTV %>%
  filter(iter == first_hit_wass)

```



